---
title: "HW 1"
author: "Kevin Ahn"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Question 2.8

### a.

$z_{[.05]} = P(Z \ge \alpha) = 0.05$
$P(Z < \alpha) = 0.95$

From Appendix A

$P(Z < 1.64) = 0.9495$ $P(Z < 1.65) = 0.9505$
$\alpha$ is appromixately 1.645

From R

```{r 28a}
qnorm(0.95)
```

### b. 

$z_{[.02]} = P(Z \ge \alpha) = 0.02$
$P(Z < \alpha) = 0.98$

From Appendix A

$P(Z < 2.05) = 0.9798$ $P(Z < 2.06) = 0.9803$
$\alpha$ is appromixately 2.055

From R
```{r 28b}
qnorm(0.98)
```

### c. 

$z_{[.01]} = P(Z \ge \alpha) = 0.01$
$P(Z < \alpha) = 0.99$

From Appendix A

$P(Z < 2.32) = 0.9898$ $P(Z < 2.33) = 0.9901$
$\alpha$ is appromixately 2.327

From R
```{r 28c}
qnorm(0.99)
```

### d. 

$z_{[.0055]} = P(Z \ge \alpha) = 0.005$
$P(Z < \alpha) = 0.995$

From Appendix A

$P(Z < 2.57) = 0.9949$ $P(Z < 2.58) = 0.9951$
$\alpha$ is appromixately 2.575

From R
```{r 28d}
qnorm(0.995)
```

# Question 2.9

### a.

From Appendix A

$t_{[0.05]}^{(7)} = 1.895$

From R

```{r 29a}
qt(0.95, 7)
```

### b.

From Appendix A

$t_{[0.01]}^{(7)} = 2.998$

From R

```{r 29b}
qt(0.99, 7)
```

### c. 

From Appendix A

$t_{[0.005]}^{(7)} = 3.499$

From R

```{r 29c}
qt(0.995, 7)
```

# Question 2.10

### a.

From Appendix A

$F_{[0.05]}^{(2,5)} = 5.79$ 

From R
```{r 210a}
qf(0.95, 2, 5)
```

### b.

From Appendix A

$F_{[0.05]}^{(5,2)} = 19.30$ 

From R

```{r 210b}
qf(0.95, 5, 2)
```

# Question 2.11

### a.

From Appendix A

$\chi^2_{[0.05]}(3) = 7.815$

From R

```{r 211a}
qchisq(0.95, 3)
```

### b.

From Appendix A

$\chi^2_{[0.01]}(2) = 9.21$

From R

```{r 211b}
qchisq(0.99, 2)
```

# Question 2.12

### a. 

$P(X < 250,000) = P(Z < \frac{250,000 - 300,000}{20,000}) = P(Z < -2.5)$

```{r 212a}
pnorm(-2.5)
```

### b.

$P(260,000 \le X \le 330,000) = P(\frac{260,000 - 300,000}{20,000} \le Z \le \frac{330,000 - 300,000}{20,000}) = P(-2 \le Z \le 1.5)$

```{r 212b}
pnorm(1.5) - pnorm(-2)
```

### c. 

$P(X > 346,000) = P(Z > \frac{346,000 - 300,000}{20,000}) = P(Z > 2.3) = 1 - P(Z \le 2.3)$

```{r 212c}
1 - pnorm(2.3)
```

# Question 2.16

### a. 

For 90% confidence interval
```{r 216a 1}
n = 81
y_bar = 57.8
s = 6.02
confidence_interval = 0.9
standard_error = s / sqrt(n)
critical = qnorm((1 + confidence_interval) / 2)
error = critical * standard_error
interval = c(y_bar - error, y_bar + error)
print(interval)
```

For 95% confidence interval
```{r 216a 2}
n = 81
y_bar = 57.8
s = 6.02
confidence_interval = 0.95
standard_error = s / sqrt(n)
critical = qnorm((1 + confidence_interval) / 2)
error = critical * standard_error
interval = c(y_bar - error, y_bar + error)
print(interval)
```

For 98% confidence interval
```{r 216a 3}
n = 81
y_bar = 57.8
s = 6.02
confidence_interval = 0.98
standard_error = s / sqrt(n)
critical = qnorm((1 + confidence_interval) / 2)
error = critical * standard_error
interval = c(y_bar - error, y_bar + error)
print(interval)
```

For 99% confidence interval
```{r 216a 4}
n = 81
y_bar = 57.8
s = 6.02
confidence_interval = 0.99
standard_error = s / sqrt(n)
critical = qnorm((1 + confidence_interval) / 2)
error = critical * standard_error
interval = c(y_bar - error, y_bar + error)
print(interval)
```

### b. 

The 95% confidence interval is (56.489, 59.111). As the highest value within the interval is less than 60, National Motors can be 95% confident that the $\mu$ is less than 60 feet.

### c.

The 98% confidence interval is (56.24393, 59.35607). As the highest value within the interval is less than 60, National Motors can be 98% confident that the $\mu$ is less than 60 feet.

# Question 2.17

### a.

$H_0: \mu = 16$
$H_1: \mu \ne 16$
$n = 6, \bar y = 15.7665, s = 0.1524$

$t_0 = \frac{\bar y - H_0}{s / \sqrt{n}} = \frac{15.7665 - 16}{0.1524 / \sqrt{6}} = \frac{-0.2335}{0.06221} = -3.753$
$t_{[0.025]}^{(5)} = 2.571$
$|-3.753| > 2.571$
$\therefore$ Reject H_0
The process should be readjusted.

### b.

$t_0 = \frac{\bar y - H_0}{s / \sqrt{n}} = \frac{15.7665 - 16}{0.1524 / \sqrt{6}} = \frac{-0.2335}{0.06221} = -3.753$
$t_{[0.005]}^{(5)} = 4.032$
$|-3.753| < 4.032$
$\therefore$ Fail to reject H_0
The process shouldn't be readjusted.

# Question 2.20

### a. 

$H_0: \mu = 3.5$
$H_1: \mu > 3.5$

### b.

```{r 220b}
p = c(7, 4, 6, 7, 5, 4, 9)
mean(p)
sd(p)
```

$\bar x = 6, s = 1.825742$
$t_0 = \frac{\bar x - 3.5}{s / \sqrt{n}} = \frac{6 - 3.5}{1.825742 / \sqrt{7}} = \frac{2.5}{0.69} = 3.6$
$t_{[0.005]}^{(6)} = 3.707$
$3.6 < 3.707$
$\therefore$ Fail to reject H_0
This means that the Federal banking official's claim that the mean bad debt ratio for Ohio banks is higher than the mean for all federally insured banks is not reasonable.