---
title: "HW 2"
author: "Kevin Ahn"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Question 3.12

### a. 

```{r scatterplot}
data = read.table("Table3-9.txt", header = TRUE)
colnames(data) = c("cost", "bsize")
plot(data$bsize, data$cost, main = "Direct Labor Cost vs. Batch Size", 
     xlab = "Batch Size", ylab = "Direct Labor Cost", pch = 19)
```

### b. 

The scatter plot suggests a simple linear regression model relating y to x, as the scatter plot presents an even "rise over run", or a linear slope.

# Question 3.14

### a. 

```{r 314a}
data = read.table("Table3-9.txt", header = TRUE)
colnames(data) = c("cost", "bsize")
x = data$bsize
y = data$cost

n = length(x)
xSum = sum(x)
ySum = sum(y)
xySum = sum(x*y)
x2Sum = sum(x^2)

b1 = (n*xySum - xSum*ySum) / (n*x2Sum - (xSum^2))
b0 = mean(y) - b1*mean(x)

b0
b1
```

### b.

$b_0$, the y intercept, represents the direct labor cost when the batch size is 0. According to the data in Table 3-9, this value is $18.4875

$b_1$, the slope, represents the increase in direct labor cost per increase of batch size. According to the data in Table 3-9, this value is $10.1462

The interpretation of $b_0$ doesn't make practical sense because there cannot be a direct labor cost when the batch size is 0.

### c.

$$
\hat{y}=b_0+b_1x
$$

### d. 

$$
\hat{y}=18.4875 + 10.14626 * 60 = 627.2631
$$

# Question 3.15

### a. 

$$
\begin{aligned}
s^2 = \frac{SSE}{n - 2}\\
n = 12, SSE = 747\\
s^2 = \frac{747}{12-2} = \frac{747}{10}=74.7\\
s =\sqrt{74.7} = 8.6429
\end{aligned}
$$

### b. 

$$
\begin{aligned}
SSE = \sum(y-\hat{y})^2\\
\hat{y} = 18.4875+10.14626x\\
e_i = y_i-\hat{y}_i\\
SSE = \sum e_i^2\\
SSE = 746.7624
\end{aligned}
$$

```{r 315b}
b0 = 18.4875
b1 = 10.14626
y_hat = b0 + b1*x
SSE = sum((y - y_hat)^2)

SSE
```

# Question 3.19

```{r 319}
data = read.table("Table3-9.txt", header = TRUE)
colnames(data) = c("cost", "bsize")
x = data$bsize
y = data$cost
model = lm(y ~ x)
summary(model)
```

### a. 

$$
b_0 = 18.48751\\
b_1 = 10.14626
$$

### b.

$$
SSE = 746.7624\\
s^2 = 74.67624\\
s = 8.641541
$$

### c. 

$$
\begin{aligned}
t = \frac{b_1}{S_{b_1}}\\
b_1 = 10.14626\\
S_{b_1} = 0.08662\\
t = \frac{10.14626}{0.08662} = 117.135304
\end{aligned}
$$

### d. 

$$
H_0: \beta_1 = 0\\
H_a: \beta_1 \ne q_0\\
\alpha = 0.05\\
$$

```{r 319d}
alpha = 0.05
t = qt((0.05/2), df = 10)
t
```

$H_0$ is rejected if $|t| > t^{n - 2}_{\alpha / 2}$

$$
\begin{aligned}
|t| = 117.1353\\
t^{10}_{\frac{0.05}{2}} = -1.406095\\
|t| > t^{10}_{0.025}
\end{aligned}
$$

Null Hypothesis $H_0$ is rejected. Because $\beta_1$ isn't equal to 0, this shows that changes in x are related to the changes in the expected value of y.

### f. 

$$
\begin{aligned}
p_{value} = 2 * P(t_{10} > |t_0|)\\
p_{value} = 2 * P(t_{10} > 117.1353) = 0\\
p_{value} < \alpha=0.10\\
p_{value} < \alpha=0.05\\
p_{value} < \alpha=0.01\\
p_{value} < \alpha=0.001
\end{aligned}
$$

The p-value is much smaller than all the given $\alpha$ values. This means that we would reject the null hypothesis, $H_0$, in all cases provided. Rejecting all the cases concludes that the linear relationship between y and x is very strong.

### g.

$$
\begin{aligned}
CI = b_1 \pm t^{n-2}_{[\alpha/2]} * S_{b_1}\\
lowerBound = 10.14626 - 2.228139 * 0.08662\\
upperBound = 10.14626 + 2.228139 * 0.08662\\
CI = (9.9532,10.3392)
\end{aligned}
$$

Through this interval, we can be 95% confident that as x increases by 1, the amount that y increases will be within the interval of (9.9532,10.3392).

### i.

$$
\begin{aligned}
S_{b_0} = 4.67658\\
b_0 = 18.48751\\
t = \frac{b_0}{S_{b_0}}\\
t = \frac{18.48751}{4.67658}\\
t = 3.953212
\end{aligned}
$$

### k. 

$$
\begin{aligned}
SS_{xx} = \sum (x_i - \bar{x})^2\\
SS_{xx} = (5 - 45.66667)^2 + (62 - 45.66667)^2 + ... + (75 - 45.66667)^2\\
SS_{xx} = 9952.667\\
S_{b_1} = \frac{s}{\sqrt{SS_{xx}}}\\
S_{b_1} = \frac{8.642}{\sqrt{9952.667}}\\
S_{b_1} = 0.08662525
\end{aligned}
$$

## Question 2

$$SSTO = \sum_{i = 1}^n (y_i - \bar y)^2 = \sum_{i = 1}^n (y_i^2 + \bar y^2 - 2 * y_i * \bar y) $$

$$SSTO = \sum_{i = 1}^n y_i^2 + \sum_{i = 1}^n \bar y^2 - \sum_{i = 1}^n 2 * y_i * \bar y $$
$$SSTO = \sum_{i = 1}^n y_i^2 + \sum_{i = 1}^n \bar y^2 - 2y * \sum_{i = 1}^ny_i $$
$$SSTO = \sum_{i = 1}^n y_i^2 + n * \bar y^2 - 2 * \bar y^2 * n $$
$$SSTO = \sum_{i = 1}^n y_i^2 - n * \bar y^2 $$
$$\therefore \sum_{i = 1}^n (y_i - \bar y)^2 = \sum_{i = 1}^n y_i^2 - n * \bar y^2$$