---
title: "HW 3"
author: "Kevin Ahn"
date: '`r Sys.Date()`'
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Question 3.23

```{r 323a}
data = read.table("Table3-9.txt", header = TRUE)
colnames(data) = c("cost", "bsize")
x = data$bsize
y = data$cost
xBar = mean(x)
SSxx = sum((x - xBar)^2)
model = lm(y ~ x)
summary(model)
```

### a.

Point estimate

$$
\begin{aligned}
\hat{y} = b_0 + b_1 * 60\\
\hat{y} = 18.48751 + 10.14626*60\\
\hat{y} = 627.2631
\end{aligned}
$$
95% Confidence Interval

$$
\begin{aligned}
CI = \hat{y} \pm t_{\alpha / 2}^{(n - 2)} * s * \sqrt{DV}\\
DV = \frac{1}{n} + \frac{(x_0 - \bar{x})^2}{SSxx}\\
DV = \frac{1}{12} + \frac{(60 - 45.666)}{9952.667}\\
DV = 0.1039755\\
CI = 627.2631 \pm t_{0.025}^{10} * 8.642 * \sqrt{0.1039755}\\
CI = (621.0541, 633.4721)
\end{aligned}
$$


### b.

Point estimate

$$
\begin{aligned}
\hat{y} = b_0 + b_1 * 60\\
\hat{y} = 18.48751 + 10.14626*60\\
\hat{y} = 627.2631
\end{aligned}
$$
95% Prediction Interval

$$
\begin{aligned}
PI = \hat{y} \pm t_{\alpha / 2}^{(n - 2)} * s * \sqrt{1 + DV}\\
PI = 627.2631 \pm t_{0.025}^{10} * 8.642 * \sqrt{1.1039755}\\
PI = (607.0312, 647.495)
\end{aligned}
$$

### c.

For 99% Confidence and Prediction Intervals

$$
\begin{aligned}
CI = 627.2631 \pm t_{0.005}^{10} * 8.642 * \sqrt{0.1039755}\\
CI = (618.4315, 636.0947)\\
PI = 627.2631 \pm t_{0.005}^{10} * 8.642 * \sqrt{1.1039755}\\
PI = (598.4856, 656.0406)
\end{aligned}
$$

# Question 3.27

### a. 

$$
\begin{aligned}
SSTO = SSR + SSE\\
SSR = \sum (\hat{y_i} - \bar y)^2\\
SSE = \sum (y_i - \hat{y_i})^2 
\end{aligned}
$$
```{r 327a}
yBar = mean(y)
SSTO = sum((y - yBar)^2)
rSquared = summary(model)$r.squared
SSR = rSquared * SSTO
SSE = SSTO - SSR
r = sqrt(rSquared)
```

$$
\begin{aligned}
SSTO = 1025340\\
SSR = 1024593\\
SSE = 746.7624\\
r^2 = 0.9992717\\
r = 0.9996358
\end{aligned}
$$

# Question 3.31

### a.

$$
F = \frac{SSR}{SSE/(n - 2)}
$$
```{r 331a}
yHat = predict(model)
residual = y - yHat
TSS = sum((y - mean(y))^2)
SSE = sum(residual^2)
SSR = TSS - SSE
n = length(y)
F = SSR / (SSE / (n - 2))
```

F-statistic: 13720.472 on 1 and 10 Degrees of Freedom

### c.

$$
\begin{aligned}
H_0: \beta_1 = 0 \\
H_a: \beta_1 \ne0\\
\alpha = 0.01
\end{aligned}
$$
We reject if $F(model) > F_{\alpha}$

```{r 331c}
qf(1 - 0.01, 1, 10)
```

$$
\begin{aligned}
F(model) = 13720\\
F_\alpha = 10.04429\\
F(model) > F_{\alpha}
\end{aligned}
$$
We reject the null hypothesis as $F(model)$ is greater than $F_{\alpha}$. This suggests that there's a significant relationship between x and y.

### d.

$$
\begin{aligned}
H_0: \beta_1 = 0\\
H_a: \beta_1 \ne0\\
\end{aligned}
$$
We reject if $p_{value} < \alpha$

$$
\begin{aligned}
p_{value}= P(F^{(1, 10)} > 13720)\\
p_{value} < \alpha=0.10\\
p_{value} < \alpha=0.05\\
p_{value} < \alpha=0.01\\
p_{value} < \alpha=0.001
\end{aligned}
$$
$p_{value}$ is less than all significance levels. We reject the null hypothesis.

### e.

$$
\begin{aligned}
t_0 = \frac{b_1}{S_{b_1}}\\
S_{b_1} = \frac{s}{\sqrt{SSxx}}\\
t_0 = \frac{10.14626}{0.08662}\\
{t_0}^2 = 13720.68=F(model)
\end{aligned}
$$

$F_{[.05]}$ rejection point is the square of $t_{.025}^{n - 2}$

```{r 331e}
t = qt(0.025, 10, lower.tail = FALSE)
f = qf(0.05, 1, 10, lower.tail = FALSE)
t
t^2
f
```

# Question 2

$$
\begin{aligned}
y_i = \beta_0 + \epsilon_i\\
\epsilon_i = y_i - \beta_0\\
L = \sum_{i = 1}^n \epsilon_i^2\\
L = \sum_{i = 1}^n (y_i - \beta_0)^2\\
\frac{dL}{d\beta_0} = \sum_{i = 1}^n -2(y_i - \beta_0)\\
\sum_{i = 1}^n -2(y_i - \beta_0) = 0\\
\sum_{i = 1}^n y_i  = n*\beta_0\\
n * \bar y = n * \beta_0\\
\beta_0 = \bar y\\
\hat \beta_0 = \bar y
\end{aligned}
$$

# Question 3

### a. 

```{r 3a}
a = c(2, 4, 6, 8, 10, 12, 14, 16, 18, 20, 22, 24)
A = matrix(a, nrow = 3, byrow = TRUE)
A
```

### b.

```{r}
b = c(3, 4, 2, 0, 0, 0, 0, 9, 3, 3, 2, 1, 0, 1, 3, 9)
B = matrix(b, nrow = 4,byrow = TRUE)
B
```

### c.

```{r}
bInverse = solve(B)
answer = A %*% bInverse
answer
```

$$
AB^{-1}=
\begin{bmatrix}
 0.5555556& -0.6419753& -0.2222222& 1.444444\\
-0.3333333& -1.0370370&  3.3333333& 2.333333\\
-1.2222222& -1.4320988&  6.8888889& 3.222222
\end{bmatrix}_{4x4}
$$

### d.

```{r}
answer = rep(5:9, c(3, 0, 3, 3, 2))
answer
```